---
title: "Case Study One: Social behaviour during a badger lifetime"
author: "Matthew Silk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CaseStudyOne}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = 'center'
  
)
```

#### Example of using CMRnet to generate movement networks of European badgers for use in analyses linking deaths caused by collisions with cars to the position of groups within movement networks

***

#### Background

This case study is based on a long-term CMR dataset from a high-density population of European badgers in the United Kingdom (Delahay et al., 2013; McDonald, Robertson, & Silk, 2018). Badgers in medium to high-density populations live in territorial social groups centred on a main sett (refuge/burrow system), resulting in modular social networks in which they interact frequently with other group members but much less often with individuals from other groups (Weber et al., 2013). The study population is trap-sampled four times per year (in May-June, July-August, September-October and December-January) using baited cage traps deployed at all know active sett locations (see  McDonald et al., 2018 for more detail). Badgers give birth in late-winter (the date of birth of all cubs is assumed to be 1st February) and no trapping takes place while females have dependent cubs.

Here we use the co-capture network construction features of CMRnet to examine how the social network position of badgers changes during their lifetime. Individuals that are more strongly connected to those in other groups will be more central to the population network. We expect that the centrality of individuals in social networks peaks when levels of reproductive behaviour as this drives many between-group movements and interactions. Badgers in our study population demonstrate body mass and immune senescence, and so we would expect network centrality to peak in early adulthood, subsequent to maturity but prior to condition beginning to decline. We would also predict that this peak will be higher in males than females, due to differences in social and spatial behaviour.

***

#### Setup and data manipulation

First we set up our environment.

```{r load_packages, message=FALSE, warning=FALSE}
# load packages
library(lubridate)
library(CMRnet)
library(glmmTMB)
library(igraph)

```

The data we use here is not documented in the package, but can be accessed to recreate this analyses by using `CMRnet:::capturesData` and `CMRnet:::individualData`. The data is on ... 

```{r load_data, echo=FALSE}
captures <- CMRnet:::capturesData
indivdat <- CMRnet:::individualData
```

```{r show_data}
# look at dimensions of data
str(captures)
str(indivdat)
```

We then do some data manipulation to fill in data on the age and sex of badgers in the captures data set. We also wrangle the date formats and create a dataframe in the format needed for **CMRnet**. 

```{r data_manipulation}
# create empty columns for sex and age
captures$sex<-rep(NA,nrow(captures))
captures$age<-rep("ADULT",nrow(captures))

# fill in data on the age and sex of badgers in the captures dataset
for(i in 1:nrow(captures)){
  if(captures$tattoo[i]%in%indivdat$tattoo){
    captures$sex[i]<-as.character(indivdat$sex[indivdat$tattoo==as.character(captures$tattoo[i])])
    if(substr(captures$date[i],7,10)==indivdat$year_fc[indivdat$tattoo==as.character(captures$tattoo[i])]&(is.na(indivdat$age_fc[indivdat$tattoo==as.character(captures$tattoo[i])])==FALSE)&indivdat$age_fc[indivdat$tattoo==as.character(captures$tattoo[i])]=="CUB"){
      captures$age[i]<-"CUB"
    }
    if(substr(captures$date[i],7,10)==indivdat$year_fc[indivdat$tattoo==as.character(captures$tattoo[i])]&(is.na(indivdat$age_fc[indivdat$tattoo==as.character(captures$tattoo[i])])==TRUE)){
      captures$age[i]<-NA
    }
  }
}

# Convert sex to a factor
captures$sex <- as.factor(captures$sex)

# Add a date column in correct format
captures$date2 <- as.Date(paste0(substr(captures$date,7,10),"-",substr(captures$date,4,5),"-",substr(captures$date,1,2)))

# Add year, month and day columns
captures$year <- lubridate::year(captures$date2)
captures$month <- lubridate::month(captures$date2)
captures$day <- lubridate::day(captures$date2)

# Create a year of study column
captures$year_of_study <- rep(NA,nrow(captures))
for(i in 1:nrow(captures)){
  if(captures$month[i]>1){
    captures$year_of_study[i]<-as.numeric(captures$year[i])-1985
  }
  if(captures$month[i]==1){
    captures$year_of_study[i]<-as.numeric(captures$year[i])-1986
  }  
}

# Subset data to only include 30 years - Feb 1986 to Jan 2016
captures2 <- captures[captures$year_of_study>0&captures$year_of_study<31,]

# Create a dataframe in the correct format for CMRnet
captures3<-data.frame(captures2$tattoo,captures2$socg,rep(NA,nrow(captures2)),rep(NA,nrow(captures2)),captures2$date2)
names(captures3)<-c("id","loc","x","y","date")

# Assign fake coordinates to locations using there factor ids as they are not used anyway in this analysis
captures3$x <- as.numeric(as.factor(captures3$loc))
captures3$y <- as.numeric(as.factor(captures3$loc))

str(captures3)

```

***

#### Create social networks

We now create the social networks using `DynamicNetCreate` because we have data on co-capture of badgers across the thirty year period.

-We choose a 12 month network window that runfs from February-January to correspond with badger life-history (cubs are born in later winter) and trapping strategy.
-We choose a 30 day interaction window. As trapping is (approximately) quarterly this ensures that all individuals captured in the same trapping event at the same location are considered to be co-captured.
-We set the space window to zero so that individuals have to be captured at the same location to be co-captured.

```{r network_creation, message=FALSE, results="hide"}

#Define parameters for first time period
mindate<-"1986-02-01"
maxdate<-"1987-02-01"
intwindow<-30
netwindow<-12
overlap<-0
spacewindow<-0

#Create network for first time period
netdat1<-DynamicNetCreate(data=captures3,
                          intwindow=intwindow,
                          mindate=mindate,
                          maxdate=maxdate, 
                          netwindow=netwindow, 
                          overlap=overlap, 
                          spacewindow=spacewindow)

#Create lists to store network data
netdats<-list()
netdats[[1]]<-netdat1

#Loop to create networks for remaining time periods
for(i in seq(1987,2015,1)){
  mindate<-paste0(i,"-02-01")
  maxdate<-paste0(i+1,"-02-01")
  intwindow<-30
  netwindow<-12
  overlap<-0
  spacewindow<-0
  netdats[[i-1985]]<-DynamicNetCreate(data=captures3,  
                                      intwindow=intwindow, 
                                      mindate=mindate, 
                                      maxdate=maxdate, 
                                      netwindow=netwindow, 
                                      overlap=overlap, 
                                      spacewindow=spacewindow)

}
```
#### Calculate network metrics

We then calculate network metrics on the observed dataset. The networks calculated are:

- Degree
- Eigenvector Centrality
- Betweenness
- PageRank Centrality

In this analysis we use (unweighted) Degree.

```{r calculate_network_metrics}

# Create list to store results
t_net_mets <- list()

# Calculate network metrics to use
for(i in 1:length(netdats)){

  inet<-graph.adjacency(netdats[[i]][[2]][,,1],weighted=NULL,mode="undirected")
  ind<-vertex_attr(inet)$name
  tp<-rep(i,length(vertex_attr(inet)$name))

  t_net_mets[[i]]<-data.frame(ind,tp)
  t_net_mets[[i]]$deg<-igraph::degree(inet)
  t_net_mets[[i]]$eig<-igraph::eigen_centrality(inet)$vector
  t_net_mets[[i]]$bet<-igraph::betweenness(inet)
  t_net_mets[[i]]$pr<-igraph::page.rank(inet)$vector

}

# Collate results
net_mets <- do.call("rbind", t_net_mets)

# Create overall results dataframe
# Add info on age, sex and group id
net_mets$group <- rep(NA,nrow(net_mets))
net_mets$sex <- rep(NA,nrow(net_mets))
net_mets$age <- rep(NA,nrow(net_mets))
for(i in 1:nrow(net_mets)){
  net_mets$sex[i] <- unique(as.character(captures2$sex[captures2$tattoo == as.character(net_mets$ind[i])]))
  if(length(indivdat$age_fc[indivdat$tattoo == as.character(net_mets$ind[i])]) > 0){
  if(is.na(indivdat$age_fc[indivdat$tattoo == as.character(net_mets$ind[i])]) == FALSE & indivdat$age_fc[indivdat$tattoo == as.character(net_mets$ind[i])] == "CUB"){
    net_mets$age[i] <- net_mets$tp[i]-(indivdat$year_fc[indivdat$tattoo == as.character(net_mets$ind[i])]-1985)
  }
  }
  t.gr <- captures3[captures3$id==as.character(net_mets$ind[i]),]
  t.gr2 <- table(t.gr$loc)
  if(length(names(which.max(t.gr2))) == 1){
   net_mets$group[i] <- names(which.max(t.gr2))
  }
}

#Remove captures with NAs for any of these additional variables
net_mets2 <- net_mets[!is.na(net_mets$age),]
net_mets2 <- net_mets2[!is.na(net_mets2$sex),]
net_mets2 <- net_mets2[!is.na(net_mets2$group),]

#And remove individuals with erroneous ages below 0 
net_mets2 <- net_mets2[net_mets2$age>-1,]

#Create age squared variable
net_mets2$agesq <- net_mets2$age^2

#Remove all individuals only recorded once from the analysis
net_mets3 <- net_mets2[net_mets2$ind%in%names(which(table(net_mets$ind)>1)),]

#Create binary measure of whether individuals have a betweenness or not
net_mets3$betbin <- sign(net_mets3$bet)


```

#### Model the observed dataset

We now conduct the statistical modelling of the observed dataset. This is done using glmmTMB. We first fit a generalised linear mixed effects model using a negative binomial error distribution. Degree is the response variable. We have fixed effects of sex, age and age squared, and consider interactions between sex and age and sex and age-squared. We have random effects of individual, capture group and network window. We confirm that parameters esttimates from a linear mixed-effects model after a log-normal transformation are very similar, and so use these in subsequent analyses (due to computational limitations when using permutation-based methods).

```{r model_observed_data}
# first model
mod_deg2<-glmmTMB(deg~sex*age+sex*I(age^2)+(1|ind)+(1|group)+(1|tp),
                  data=net_mets3[net_mets3$age<11,],
                  family=nbinom2)

summary(mod_deg2)

# second model
mod_deg2alt<-glmmTMB(log(deg+1)~sex*age+sex*I(age^2)+(1|ind)+(1|group)+(1|tp),
                     data=net_mets3[net_mets3$age<11,],
                     family=gaussian)

summary(mod_deg2alt)

```

#### Create permuted datasets

We now create permuted datasets. These are randomisations of the dataset to see what a random network would look like.

```{r create_perms, eval=FALSE}

# First we create a list of matrices to use to restrict the swaps conducted

restricts <- list()

#Loop to create networks for remaining time periods
for(i in seq(1986,2015,1)){
  
  restricts[[i-1985]] <- array(0,dim=c(nrow(netdats[[i-1985]][[2]]),ncol(netdats[[i-1985]][[2]]),1))
  rownames(restricts[[i-1985]]) <- rownames(netdats[[i-1985]][[2]])
  colnames(restricts[[i-1985]]) <- colnames(netdats[[i-1985]][[2]])
  
  mindate <- paste0(i,"-02-01")
  maxdate <- paste0(i+1,"-02-01")
  D <- captures3
  
  names(D) <- c("id","loc","x","y","date")
  Jdays <- julian(as.Date(D$date),origin=as.Date("1970-01-01"))
  D <- data.frame(D,Jdays)
  D <- D[order(D$Jdays, D$loc,D$id),]
  D$id <- as.factor(D$id)
  D$loc <- as.factor(D$loc)
  D$x <- as.numeric(D$x)
  D$y <- as.numeric(D$y)
  start <- julian(as.Date(mindate),origin=as.Date("1970-01-01"))
  end <- julian(as.Date(maxdate),origin=as.Date("1970-01-01"))
  D2 <- D[which(D$Jdays>=start&D$Jdays<end),]
  
  tinds <- rownames(netdats[[i-1985]][[2]])
  for(ii in 1:length(tinds)){
    tlocs <- D2[D2$id==tinds[ii],]$loc
    tinds2 <- unique(D2[D2$loc%in%tlocs,]$id)
    restricts[[i-1985]][rownames(restricts[[i-1985]])%in%tinds2,tinds[ii],1]<-restricts[[i-1985]][tinds[ii],rownames(restricts[[i-1985]])%in%tinds2,1]<-1
  }
  
}

permuted_nets <- list()

for(i in 1:length(netdats)){
  permuted_nets[[i]] <- cmrRestrictedNodeswap(cmrnet=netdats[[i]],
                                              restrict=restricts[[i]],
                                              n.rand=1000, 
                                              n.burnin=500, 
                                              n.swaps=40, 
                                              multi=FALSE)
  print(paste(i,"done"))
}

```

#### Analyse permuted datasets

We can now rebuild the dataframe analysed and refit models for the permuted datasets.

```{r model_perms, eval=FALSE}

n.rand <- 1000

fixefs <- matrix(NA,nr=n.rand,nc=6)
ranefs <- matrix(NA,nr=n.rand,nc=4)

for(j in 1:n.rand){

# Create list to store results
t_net_mets <- list()

# Calculate network metrics to use
for(i in 1:length(netdats)){

  inet <- graph.adjacency(permuted_nets[[i]][[1]][,,j],weighted=NULL,mode="undirected")
  ind <- vertex_attr(inet)$name
  tp <- rep(i,length(vertex_attr(inet)$name))

  t_net_mets[[i]] <- data.frame(ind,tp)
  t_net_mets[[i]]$deg <- igraph::degree(inet)

}

# Collate results
net_mets <- do.call("rbind", t_net_mets)

# Create overall results dataframe
# Add info on age, sex and group id
net_mets$group <- rep(NA,nrow(net_mets))
net_mets$sex <- rep(NA,nrow(net_mets))
net_mets$age <- rep(NA,nrow(net_mets))

for(i in 1:nrow(net_mets)){
  net_mets$sex[i] <- unique(as.character(captures2$sex[captures2$tattoo==as.character(net_mets$ind[i])]))
  if(length(indivdat$age_fc[indivdat$tattoo == as.character(net_mets$ind[i])])>0){
  if(is.na(indivdat$age_fc[indivdat$tattoo == as.character(net_mets$ind[i])])== FALSE&indivdat$age_fc[indivdat$tattoo == as.character(net_mets$ind[i])] == "CUB"){
    net_mets$age[i] <- net_mets$tp[i]-(indivdat$year_fc[indivdat$tattoo == as.character(net_mets$ind[i])]-1985)
  }
  }
  t.gr <- captures3[captures3$id == as.character(net_mets$ind[i]),]
  t.gr2 <- table(t.gr$loc)
  if(length(names(which.max(t.gr2))) == 1){
   net_mets$group[i] <- names(which.max(t.gr2))
  }
}

# Remove captures with NAs for any of these additional variables
net_mets2 <- net_mets[!is.na(net_mets$age),]
net_mets2 <- net_mets2[!is.na(net_mets2$sex),]
net_mets2 <- net_mets2[!is.na(net_mets2$group),]

# And remove individuals with erroneous ages below 0 
net_mets2 <- net_mets2[net_mets2$age>-1,]

# Create age squared variable
net_mets2$agesq <- net_mets2$age^2

#Remove all individuals only recorded once from the analysis
net_mets3 <- net_mets2[net_mets2$ind%in%names(which(table(net_mets$ind)>1)),]

tmod <- glmmTMB(log(deg+1)~sex*age+sex*I(age^2)+(1|ind)+(1|group)+(1|tp),
                data=net_mets3[net_mets3$age<11,],
                family=gaussian)

ranefs[j,] <- c(unlist(summary(tmod)[9]$varcor),summary(mod_deg2alt)[7]$sigma^2)
fixefs[j,] <- as.vector(unlist(fixef(tmod)))[1:6]

print(paste(j,"done"))

}

```

```{r load_perm_results, echo = FALSE}
fixefs <- CMRnet:::fixefs
ranefs <- CMRnet:::ranefs
n.rand <- 1000
```

#### Plot results

We now calculate the quantiles of the reference distribution, calculate p values and produce the plots used in the paper

```{r calc_results}
# quantiles of reference distribution
quantile(fixefs[,1],c(0.025,0.975))
quantile(fixefs[,2],c(0.025,0.975))
quantile(fixefs[,3],c(0.025,0.975))
quantile(fixefs[,4],c(0.025,0.975))
quantile(fixefs[,5],c(0.025,0.975))
quantile(fixefs[,6],c(0.025,0.975))

# p values
sum(summary(mod_deg2alt)$coefficients$cond[,1][1]<fixefs[,1])/(n.rand+1)
sum(summary(mod_deg2alt)$coefficients$cond[,1][2]<fixefs[,2])/(n.rand+1)
sum(summary(mod_deg2alt)$coefficients$cond[,1][3]<fixefs[,3])/(n.rand+1)
sum(summary(mod_deg2alt)$coefficients$cond[,1][4]<fixefs[,4])/(n.rand+1)
sum(summary(mod_deg2alt)$coefficients$cond[,1][5]<fixefs[,5])/(n.rand+1)
sum(summary(mod_deg2alt)$coefficients$cond[,1][6]<fixefs[,6])/(n.rand+1)
```

```{r plot_fixefs, fig.width = 7, fig.height = 6}
# plot of fixed effects
par(mfrow=c(3,2))
hist(fixefs[,1],col="light grey",border=NA,xlim=c(2.1,2.6),main="Intercept",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(fixef(mod_deg2alt)[[1]][1],2),y=c(0,1000),lwd=2,col="red")
hist(fixefs[,2],col="light grey",border=NA,xlim=c(0,0.2),main="Main effect of sex",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(fixef(mod_deg2alt)[[1]][2],2),y=c(0,1000),lwd=2,col="red")
hist(fixefs[,3],col="light grey",border=NA,xlim=c(-0.15,0.05),main="Main effect of age",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(fixef(mod_deg2alt)[[1]][3],2),y=c(0,1000),lwd=2,col="red")
hist(fixefs[,4],col="light grey",border=NA,xlim=c(-0.005,0.015),main="Main effect of age squared",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(fixef(mod_deg2alt)[[1]][4],2),y=c(0,1000),lwd=2,col="red")
hist(fixefs[,5],col="light grey",border=NA,xlim=c(-0.1,0),main="sex * age",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(fixef(mod_deg2alt)[[1]][5],2),y=c(0,1000),lwd=2,col="red")
hist(fixefs[,6],col="light grey",border=NA,xlim=c(-0.005,0.02),main="sex * age squared",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(fixef(mod_deg2alt)[[1]][6],2),y=c(0,1000),lwd=2,col="red")
```

```{r plot_random_effs, fig.width=6, fig.height = 2}
# plot of random effects
par(mfrow=c(1,3),mar=c(5,5,3,3))
hist(ranefs[,1],col="light grey",border=NA,xlim=c(0,0.15),main="Individual",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(unlist(summary(mod_deg2alt)[9]$varcor)[1],2),y=c(0,1000),lwd=2,col="red")
hist(ranefs[,2],col="light grey",border=NA,xlim=c(0,0.15),main="Group",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(unlist(summary(mod_deg2alt)[9]$varcor)[2],2),y=c(0,1000),lwd=2,col="red")
hist(ranefs[,3],col="light grey",border=NA,xlim=c(0,0.15),main="Time Period",xlab="Estimate",cex.lab=1.3,cex.main=1.5)
lines(x=rep(unlist(summary(mod_deg2alt)[9]$varcor)[3],2),y=c(0,1000),lwd=2,col="red")
```

```{r plot_preds, fig.height=4, fig.width=6}
# plot of predicted change in degree with age
par(mfrow=c(1,1),mar=c(5,5,3,3))
x<-seq(0,14,0.01)
yf<-unlist(fixef(mod_deg2alt))[1]+unlist(fixef(mod_deg2alt))[3]*x+unlist(fixef(mod_deg2alt))[4]*x^2
ym<-unlist(fixef(mod_deg2alt))[1]+unlist(fixef(mod_deg2alt))[2]+(unlist(fixef(mod_deg2alt))[3]+unlist(fixef(mod_deg2alt))[5])*x+(unlist(fixef(mod_deg2alt))[4]+unlist(fixef(mod_deg2alt))[6])*x^2
plot(x,exp(yf)+1,col="dodger blue",type="l",lwd=4,ylab="Predicted Degree",xlab="Age",cex.lab=1.5,cex.axis=1.3,las=1,ylim=c(5,15))
lines(x,exp(ym)+1,col="firebrick",lwd=4)

for(i in 1:50){
  yf<-fixefs[i,1]+fixefs[i,3]*x+fixefs[i,4]*x^2
  ym<-fixefs[i,1]+fixefs[i,2]+(fixefs[i,3]+fixefs[i,5])*x+(fixefs[i,4]+fixefs[i,6])*x^2
  lines(x,exp(yf)+1,col=adjustcolor("dodgerblue",0.05))
  lines(x,exp(ym)+1,col=adjustcolor("firebrick",0.05))
}

```


***

#### References

- Delahay, R. J., Walker, N., Smith, G. S., Wilkinson, D., Clifton-Hadley, R. S., Cheeseman, C. L., … Chambers, M. A. (2013). Long-term temporal trends and estimated transmission rates for Mycobacterium bovis infection in an undisturbed high-density badger (Meles meles) population. Epidemiology and Infection, 141(07), 1445–1456.
- McDonald, J. L., Robertson, A., & Silk, M. J. (2018). Wildlife disease ecology from the individual to the population: Insights from a long-term study of a naturally infected European badger population. Journal of Animal Ecology, 87(1). doi:10.1111/1365-2656.12743
- Weber, N., Carter, S. P., Dall, S. R. X., Delahay, R. J., McDonald, J. L., Bearhop, S., & McDonald, R. A. (2013). Badger social networks correlate with tuberculosis infection. Current Biology, 23(20), R915–R916.

